---
title: "stat_software"
author: "ggplot2 group3"
date: "2021/6/25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse) # general
library(ggalt) # dumbbell plots
library(countrycode) # continent
library(rworldmap) # quick country-level heat maps
library(gridExtra) # plots
library(broom) # significant trends within countries
library(gganimate)
theme_set(theme_light())
# 1) Import & data cleaning

data <- read_csv("D:/stat_soft/master.csv") 


data <- data %>% 
  select(-c(`HDI for year`, `suicides/100k pop`)) %>%
  rename(gdp_for_year = `gdp_for_year ($)`, 
         gdp_per_capita = `gdp_per_capita ($)`, 
         country_year = `country-year`) %>%
  as.data.frame()
```

## R Markdown

### data


```{r,echo=FALSE}
# 2) OTHER ISSUES

# a) this SHOULD give 12 rows for every county-year combination (6 age bands * 2 genders):

# data %>% 
#   group_by(country_year) %>%
#   count() %>%
#   filter(n != 12) # note: there appears to be an issue with 2016 data
# not only are there few countries with data, but those that do have data are incomplete

data <- data %>%
  filter(year != 2016) %>% # I therefore exclude 2016 data
  select(-country_year)


# b) excluding countries with <= 3 years of data:

minimum_years <- data %>%
  group_by(country) %>%
  summarize(rows = n(), 
            years = rows / 12) %>%
  arrange(years)

data <- data %>%
  filter(!(country %in% head(minimum_years$country, 7)))


# no other major data issues found yet



# 3) TIDYING DATAFRAME
data$age <- gsub(" years", "", data$age)
data$sex <- ifelse(data$sex == "male", "Male", "Female")


# getting continent data:
data$continent <- countrycode(sourcevar = data[, "country"],
                              origin = "country.name",
                              destination = "continent")

# Nominal factors
data_nominal <- c('country', 'sex', 'continent')
data[data_nominal] <- lapply(data[data_nominal], function(x){factor(x)})


# Making age ordinal
data$age <- factor(data$age, 
                   ordered = T, 
                   levels = c("5-14",
                              "15-24", 
                              "25-34", 
                              "35-54", 
                              "55-74", 
                              "75+"))

# Making generation ordinal
data$generation <- factor(data$generation, 
                          ordered = T, 
                          levels = c("G.I. Generation", 
                                     "Silent",
                                     "Boomers", 
                                     "Generation X", 
                                     "Millenials", 
                                     "Generation Z"))

data <- as_tibble(data)


# the global rate over the time period will be useful:

global_average <- (sum(as.numeric(data$suicides_no)) / sum(as.numeric(data$population))) * 100000

# view the finalized data
glimpse(data)
```

### Global suicide rate

```{r}
data %>%
 group_by(year) %>%
 summarize(population = sum(population), 
 suicides = sum(suicides_no), 
 suicides_per_100k = (suicides / population) * 100000) %>%
 ggplot(aes(x = year, y = suicides_per_100k)) + 
 geom_line(col = "deepskyblue3", size = 1) + 
 geom_point(col = "deepskyblue3", size = 2) + 
 geom_hline(yintercept = global_average, linetype = 2, color = "grey35", size = 1) +
 labs(title = "Global Suicides (per 100k)",
 subtitle = "Trend over time, 1985 - 2015.",
 x = "Year", 
 y = "Suicides per 100k") + 
 scale_x_continuous(breaks = seq(1985, 2015, 2)) + 
 scale_y_continuous(breaks = seq(10, 20))
```

### By age

```{r}
age_plot <- data %>%
 group_by(age) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = age, y = suicide_per_100k, fill = age)) + 
 geom_bar(stat = "identity") + 
 labs(title = "Global suicides per 100k, by Age",
 x = "Age", 
 y = "Suicides per 100k") +
 scale_y_continuous(breaks = seq(0, 30, 1), minor_breaks = F)
### with time
age_time_plot <- data %>%
 group_by(year, age) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = year, y = suicide_per_100k, col = age)) + 
 geom_line() + 
 geom_point() + 
 labs(title = "Trends Over Time, by Age", 
 x = "Year", 
 y = "Suicides per 100k", 
 color = "Age") + 
 scale_x_continuous(breaks = seq(1985, 2015, 5), minor_breaks = F)
## `summarise()` has grouped output by 'year'. You can override using the `.groups` argument.
grid.arrange(age_plot, age_time_plot, ncol = 2)
```

```{r}
age_plot <- data %>%
 group_by(age) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = age, y = suicide_per_100k, fill = age)) + 
 geom_bar(stat = "identity") + 
 labs(title = "Global suicides per 100k, by Age",
 x = "Age", 
 y = "Suicides per 100k") +
 scale_y_continuous(breaks = seq(0, 30, 1), minor_breaks = F)
### with time
age_time_plot <- data %>%
 group_by(year, age) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = year, y = suicide_per_100k, col = age)) + 
 facet_grid(age ~ ., scales = "free_y") + ##分成6塊
 geom_line() + 
 geom_point() + 
 labs(title = "Trends Over Time, by Age", 
 x = "Year", 
 y = "Suicides per 100k", 
 color = "Age") + 
 scale_x_continuous(breaks = seq(1985, 2015, 5), minor_breaks = F)
## `summarise()` has grouped output by 'year'. You can override using the `.groups` argument.
grid.arrange(age_plot, age_time_plot, ncol = 2)
```

```{r}
age_plot <- data %>%
 group_by(age) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = age, y = suicide_per_100k, fill = age)) + 
 geom_bar(stat = "identity") + 
 labs(title = "Global suicides per 100k, by Age",
 x = "Age", 
 y = "Suicides per 100k") +
 theme(legend.position = "none") + ##圖例
 scale_y_continuous(breaks = seq(0, 30, 1), minor_breaks = F)
### with time
age_time_plot <- data %>%
 group_by(year, age) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = year, y = suicide_per_100k, col = age)) + 
 facet_grid(age ~ ., scales = "free_y") + ##分成6塊
 geom_line() + 
 geom_point() + 
 labs(title = "Trends Over Time, by Age", 
 x = "Year", 
 y = "Suicides per 100k", 
 color = "Age") + 
 theme(legend.position = "none") + 
 scale_x_continuous(breaks = seq(1985, 2015, 5), minor_breaks = F)
## `summarise()` has grouped output by 'year'. You can override using the `.groups` argument.
grid.arrange(age_plot, age_time_plot, ncol = 2)

```

### By sex

```{r}
sex_plot <- data %>%
group_by(sex) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = sex, y = suicide_per_100k, fill = sex)) + 
 geom_bar(stat = "identity") + 
 labs(title = "Global suicides (per 100k), by Sex",
 x = "Sex", 
 y = "Suicides per 100k") +
 theme(legend.position = "none") + 
 scale_y_continuous(breaks = seq(0, 25), minor_breaks = F)
### with time
sex_time_plot <- data %>%
 group_by(year, sex) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = year, y = suicide_per_100k, col = factor(sex))) + 
 facet_grid(sex ~ ., scales = "free_y") + 
 geom_line() + 
 geom_point() + 
 labs(title = "Trends Over Time, by Sex", 
 x = "Year", 
 y = "Suicides per 100k", 
 color = "Sex") + 
 theme(legend.position = "none") + 
 scale_x_continuous(breaks = seq(1985, 2015, 5), minor_breaks = F)
## `summarise()` has grouped output by 'year'. You can override using the `.groups` argument.
grid.arrange(sex_plot, sex_time_plot, ncol = 2)
```

```{r}
sex_plot <- data %>%
group_by(sex) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = sex, y = suicide_per_100k)) + 
 geom_bar(stat = "identity") + 
 labs(title = "Global suicides (per 100k), by Sex",
 x = "Sex", 
 y = "Suicides per 100k") +
 theme(legend.position = "none") + 
 scale_y_continuous(breaks = seq(0, 25), minor_breaks = F)
### with time
sex_time_plot <- data %>%
 group_by(year, sex) %>%
 summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) *
100000) %>%
 ggplot(aes(x = year, y = suicide_per_100k)) + 
 facet_grid(sex ~ ., scales = "free_y") + 
 geom_line() + 
 geom_point() + 
 labs(title = "Trends Over Time, by Sex", 
 x = "Year", 
 y = "Suicides per 100k", 
 color = "Sex") + 
 theme(legend.position = "none") + 
 scale_x_continuous(breaks = seq(1985, 2015, 5), minor_breaks = F)
## `summarise()` has grouped output by 'year'. You can override using the `.groups` argument.
grid.arrange(sex_plot, sex_time_plot, ncol = 2)
```

### gganimate

```{r,echo= TRUE}
a <- data %>%
  group_by(year) %>%
  summarize(population = sum(population), 
            suicides = sum(suicides_no), 
            suicides_per_100k = (suicides / population) * 100000) %>%
  ggplot(aes(x = year, y = suicides_per_100k)) + 
  geom_line(col = "deepskyblue3", size = 1) + 
  geom_point(col = "deepskyblue3", size = 2) + 
  geom_hline(yintercept = global_average, linetype = 2, color = "grey35", size = 1) +
  labs(title = "Global Suicides (per 100k)",
       subtitle = "Trend over time, 1985 - 2015.",
       x = "Year", 
       y = "Suicides per 100k") + 
  scale_x_continuous(breaks = seq(1985, 2015, 2)) + 
  scale_y_continuous(breaks = seq(10, 20))+
  transition_reveal(year)
animate(a, renderer = gifski_renderer())

```

### gganimate by continent

```{r}
continent <- data %>%
  group_by(continent) %>%
  summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) * 100000) %>%
  arrange(suicide_per_100k)

continent$continent <- factor(continent$continent, ordered = T, levels = continent$continent)

continent_time <- data %>%
  group_by(year, continent) %>%
  summarize(suicide_per_100k = (sum(as.numeric(suicides_no)) / sum(as.numeric(population))) * 100000)

continent_time$continent <- factor(continent_time$continent, ordered = T, levels = continent$continent)

continent_time_plot <- ggplot(continent_time, aes(x = year, y = suicide_per_100k, col = factor(continent))) + 
  facet_grid(continent ~ ., scales = "free_y") + 
  geom_line() + 
  geom_point() + 
  labs(title = "Trends Over Time, by Continent", 
       x = "Year", 
       y = "Suicides per 100k", 
       color = "Continent") + 
  theme(legend.position = "none", title = element_text(size = 10)) + 
  scale_x_continuous(breaks = seq(1985, 2015, 5), minor_breaks = F)+
  transition_reveal(year)
animate(continent_time_plot, renderer = gifski_renderer())
```

### Bar chart race (suicide rate)

```{r,warning=FALSE,echo=FALSE}
data %>%
  group_by(country, year,continent) %>%
  summarise_at(c("population", "suicides_no"), sum, na.rm = TRUE)%>%
  # assign ranking
  mutate(suicides_per_100k = suicides_no/population*100000)%>%
  select("country","year","suicides_per_100k","continent")%>%
  group_by(year)%>%
  arrange(year, -suicides_per_100k)%>%
  # assign ranking
  mutate(rank = 1:n())%>%
  filter(rank <= 10)->
  ranked_by_year
my_theme <- theme_classic(base_family = "Times") +
  theme(axis.text.y = element_blank()) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.line.y = element_blank()) +
  theme(legend.background = element_rect(fill = "gainsboro")) +
  theme(plot.background = element_rect(fill = "gainsboro")) +
  theme(panel.background = element_rect(fill = "gainsboro"))
ranked_by_year %>%
  ggplot() +
  aes(xmin = 0 ,
      xmax = suicides_per_100k) +
  aes(ymin = rank - .45,
      ymax = rank + .45,
      y = rank)+
  facet_wrap(~ year)+
  geom_rect(alpha = .7)+
  aes(fill = continent)+
  scale_x_continuous(
    limits = c(-20, 55),
    breaks = c(0, 15, 30, 45)) +
  geom_text(hjust = "right",
            aes(label = country),
            x = -1)+
  scale_y_reverse()+
  labs(x = 'Suicides per 100k') +
  labs(y = "")+
  my_theme ->
  my_plot
aa<- my_plot +  
  facet_null() +  
  scale_x_continuous(
    limits = c(-20, 55),
    breaks = c(0, 15, 30, 45))+
  geom_text(x = 44 , y = -10,  
            family = "Times",  
            aes(label = as.character(year)),  
            size = 15, col = "grey18") +  
  aes(group = country) +  
  transition_states(states = year,transition_length = 4,state_length = 1)+
  ease_aes('cubic-in-out')
animate(aa,300,fps = 10)
```

### data preprocessing

```{r}
data %>%
  group_by(country, year,continent) %>%
  summarise_at(c("population", "suicides_no"), sum, na.rm = TRUE)

```

### calculate suicide rate

```{r}
data %>%
  group_by(country, year,continent) %>%
  summarise_at(c("population", "suicides_no"), sum, na.rm = TRUE)%>%
  # assign ranking
  mutate(suicides_per_100k = suicides_no/population*100000)%>%
  select("country","year","suicides_per_100k","continent")
```

### groupby year and sort by year and suicide rate

```{r}
data %>%
  group_by(country, year,continent) %>%
  summarise_at(c("population", "suicides_no"), sum, na.rm = TRUE)%>%
  # assign ranking
  mutate(suicides_per_100k = suicides_no/population*100000)%>%
  select("country","year","suicides_per_100k","continent")%>%
  group_by(year)%>%
  arrange(year, -suicides_per_100k)
```

### create rank

```{r}
data %>%
  group_by(country, year,continent) %>%
  summarise_at(c("population", "suicides_no"), sum, na.rm = TRUE)%>%
  # assign ranking
  mutate(suicides_per_100k = suicides_no/population*100000)%>%
  select("country","year","suicides_per_100k","continent")%>%
  group_by(year)%>%
  arrange(year, -suicides_per_100k)%>%
  # assign ranking
  mutate(rank = 1:n())%>%
  filter(rank <= 10)
```


```{r,echo=FALSE}
country_year <- data %>%
  group_by(country, year,continent) %>%
  summarise_at(c("population", "suicides_no"), sum, na.rm = TRUE)%>%
  # assign ranking
  mutate(suicides_per_100k = suicides_no/population*100000)%>%
  select("country","year","suicides_per_100k","continent")%>%
  group_by(year)%>%
  arrange(year, -suicides_per_100k)%>%
  # assign ranking
  mutate(rank = 1:n())%>%
  filter(rank <= 10)->
  ranked_by_year
```

### decide the theme

```{r,warning=FALSE}
my_theme <- theme_classic(base_family = "Times") +
  theme(axis.text.y = element_blank()) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.line.y = element_blank()) +
  theme(legend.background = element_rect(fill = "gainsboro")) +
  theme(plot.background = element_rect(fill = "gainsboro")) +
  theme(panel.background = element_rect(fill = "gainsboro"))

```

### first layer (aes)

```{r}
ranked_by_year %>%
  ggplot() +
  aes(x = suicides_per_100k,y = rank)


```

### adjust x and y aes

```{r}
plot <- ranked_by_year %>%
  ggplot() +
  aes(xmin = 0 ,
      xmax = suicides_per_100k) +
  aes(ymin = rank - .45,
      ymax = rank + .45,
      y = rank)
plot

```

### facet by year

```{r}
plot <- plot+facet_wrap(~ year)
plot

```

### add bar

```{r}
plot <- plot+
  geom_rect(alpha = .7)
plot
```

### fill the color by continent

```{r}
plot <- plot+
  aes(fill = continent)
plot
```

### scale x

```{r}
plot <- plot+
  scale_x_continuous(
    limits = c(-20, 55),
    breaks = c(0, 15, 30, 45))
plot
```

### add text

```{r,warning=FALSE}
plot <- plot +
  geom_text(hjust = "right",
            aes(label = country),
            x = -1)
plot
```

### reverse y and add labels and theme

```{r,warning=FALSE}
myplot <- plot+
  scale_y_reverse()+
  labs(x = 'Suicides per 100k') +
  labs(y = "")+
  my_theme
```

### facet null (merge the plot from each year)

```{r,warning=FALSE}
racingbar <- my_plot +  
  facet_null() +  
  scale_x_continuous(
    limits = c(-20, 55),
    breaks = c(0, 15, 30, 45))
racingbar

```

### add the text which represents year

```{r,warning=FALSE}
racingbar <- racingbar+
  geom_text(x = 44 , y = -10,  
            family = "Times",  
            aes(label = as.character(year)),  
            size = 15, col = "grey18") +  
  aes(group = country)
racingbar

```

### convert the static plot into animated plot

```{r,warning=FALSE}
racingbar <-  racingbar +  
  transition_states(states = year,transition_length = 4,state_length = 1)+
  ease_aes('cubic-in-out')

```

### output

```{r,warning=FALSE}
animate(racingbar,300,fps = 10)
anim_save("racingbar.gif", racingbar)
```

### animate以前是施文千做的

### animate是陳詠翰做的

[gganimate](https://gganimate.com/index.html)

[reference(1)](https://evamaerey.github.io/little_flipbooks_library/racing_bars/racing_barcharts.html#1)

[ggplot2tutorial](http://r-statistics.co/Complete-Ggplot2-Tutorial-Part1-With-R-Code.html)

[dataset](https://www.kaggle.com/lmorgan95/r-suicide-rates-in-depth-stats-insights)

# Thank You !!!